Supplementary material for\\ A Benchmark of Named Entity Recognition Approaches in Historical Documents\\ Application to 19$^{th}$ Century French Directories

### 2022-01-20 src-latex/parts/50-ner-xp1-training-set-size-and-pretraining.tex
Updates from Overleaf
Commented out table 2, merged it with figure 6

The three models are fine-tuned on the NER task 5 times using each of the 8 training subsets, with an early stopping criterion based on the number of training steps without improvement of the F1-score. 
This patience threshold is set to 1600 steps for SpaCY NER and 3 evaluations (1 evaluation every 100 steps) for CmBERT and CmBERT+ptrn.
The metrics are measured for the 24 resulting NER models on the common test set and averaged over the runs.

@@

\subsection{Results and discussion}
Table \ref{tab:experiment-1-models-performances} and Figure \ref{fig:f1-vs-trainsize} display the averaged precision, recall and F1-score for all models on the 8 subsets created from the groundtruth.
CmBERT, CmBERT+ptrn and SpaCy NER display the same behaviour: performances increase dramatically with the number of training examples and rapidly reach an area of slower progress around 1000 examples.
The F1 score increases by 4.6 points between 49 and 796 examples for CmBERT (resp. 1.6 for CmBERT+ptrn and 5.1 for SpaCy NER) but only by 1 point between 796 and 6373 examples (resp. 0.6 and 1.4).
The models derived from CamemBERT always outperform the Spacy model.

@@

\begin{table}[ht!]
\centering
\caption{\label{tab:experiment-1-models-performances} F1 score, precision and recall measured on the fine-tuned models CmBERT, CmBERT+ptrn and SpaCy NER in experiment 1 for 8 training sets of increasing sizes.}
\begin{tabular}{llrrrrrrrr}
       & Training examples &  49   &  99   &  199  &  398  &  796  &  1593 &  3186 &  6373 \\
       & \% & 0.8   & 1.6   & 3.1   & 6.2   & 12.5  & 25.0  & 50.0  & 100.0 \\
\midrule\bottomrule
\multirow{3}{*}{\rotatebox{90}{F1 score}} & CmBERT &  89.5 &  90.5 &  92.7 &  93.3 &  \textbf{94.1} &  \textbf{94.9} &  \textbf{94.6} &  \textbf{95.1} \\
       & CmBERT-ptrn &  \textbf{92.2} &  \textbf{92.9} &  \textbf{93.6} &  \textbf{93.8} &  93.8 &  94.1 &  \textbf{94.6} &  94.4 \\
       & SpaCy NER &  87.0 &  89.0 &  90.3 &  91.9 &  92.1 &  92.8 &  93.2 &  93.5 \\
\cline{1-10}
\multirow{3}{*}{\rotatebox{90}{Precision}} & CmBERT &  87.4 &  88.7 &  91.5 &  92.7 &  93.3 &  94.9 &  93.9 &  95.1 \\
       & CmBERT-ptrn &  90.8 &  91.8 &  92.9 &  93.0 &  93.0 &  93.4 &  94.1 &  93.9 \\
       & SpaCy NER &  85.6 &  87.7 &  90.0 &  92.0 &  92.4 &  92.8 &  93.1 &  93.7 \\
\cline{1-10}
\multirow{3}{*}{\rotatebox{90}{Recall}} & CmBERT &  91.6 &  92.5 &  93.9 &  93.9 &  94.9 &  94.9 &  95.4 &  95.1 \\
       & CmBERT-ptrn &  93.6 &  94.0 &  94.4 &  94.6 &  94.6 &  94.8 &  95.0 &  94.9 \\
       & SpaCy NER &  88.6 &  90.4 &  90.7 &  91.7 &  91.9 &  92.8 &  93.3 &  93.4 \\
\end{tabular}
\end{table}

\begin{figure}[ht!]
	   \center{\includegraphics[width=\textwidth]
	       {images/experiment-1-models-performances.pdf}}
	  \caption{\label{fig:f1-vs-trainsize} Graph view of the metrics displayed in~\cref{tab:experiment-1-models-performances}. As SpaCy NER is outperformed by both BERT models from the start, CmBERT slightly surpasses the pre-trained version when the number of training examples becomes large.}.
\end{figure}

>>>>

The three models are fine-tuned on the NER task 5 times using each of the 8 training subsets, with an early stopping criterion based on the number of training steps without improvement of the F1-score. 
This patience threshold is set to 1600 steps for SpaCy NER and 3 evaluations (1 evaluation every 100 steps) for CmBERT and CmBERT+ptrn.
The metrics are measured for the 24 resulting NER models on the common test set and averaged over the runs.

@@

\subsection{Results and discussion}
Figure \ref{fig:f1-vs-trainsize} displays the averaged precision, recall and F1-score for all models on the 8 subsets created from the groundtruth.
CmBERT, CmBERT+ptrn and SpaCy NER display the same behaviour: performances increase dramatically with the number of training examples and rapidly reach an area of slower progress around 1000 examples.
The F1 score increases by 4.6 points between 49 and 796 examples for CmBERT (resp. 1.6 for CmBERT+ptrn and 5.1 for SpaCy NER) but only by 1 point between 796 and 6373 examples (resp. 0.6 and 1.4).
The models derived from CamemBERT always outperform the Spacy model.

@@

% MOVE THIS TABLE TO THE APPENDIX
%\begin{table}[ht!]
%\centering
%\caption{\label{tab:experiment-1-models-performances} F1 score, precision and recall measured on the fine-tuned models CmBERT, CmBERT+ptrn and SpaCy NER in experiment 1 for 8 training sets of increasing sizes.}
%\begin{tabular}{llrrrrrrrr}
%       & Training examples &  49   &  99   &  199  &  398  &  796  &  1593 &  3186 &  6373 \\
%       & \% & 0.8   & 1.6   & 3.1   & 6.2   & 12.5  & 25.0  & 50.0  & 100.0 \\
%\midrule\bottomrule
%\multirow{3}{*}{\rotatebox{90}{F1 score}} & CmBERT &  89.5 &  90.5 &  92.7 &  93.3 &  \textbf{94.1} &  \textbf{94.9} &  \textbf{94.6} &  \textbf{95.1} \\
%       & CmBERT-ptrn &  \textbf{92.2} &  \textbf{92.9} &  \textbf{93.6} &  \textbf{93.8} &  93.8 &  94.1 &  \textbf{94.6} &  94.4 \\
%       & SpaCy NER &  87.0 &  89.0 &  90.3 &  91.9 &  92.1 &  92.8 &  93.2 &  93.5 \\
%\cline{1-10}
%\multirow{3}{*}{\rotatebox{90}{Precision}} & CmBERT &  87.4 &  88.7 &  91.5 &  92.7 &  93.3 &  94.9 &  93.9 &  95.1 \\
%       & CmBERT-ptrn &  90.8 &  91.8 &  92.9 &  93.0 &  93.0 &  93.4 &  94.1 &  93.9 \\
%       & SpaCy NER &  85.6 &  87.7 &  90.0 &  92.0 &  92.4 &  92.8 &  93.1 &  93.7 \\
%\cline{1-10}
%\multirow{3}{*}{\rotatebox{90}{Recall}} & CmBERT &  91.6 &  92.5 &  93.9 &  93.9 &  94.9 &  94.9 &  95.4 &  95.1 \\
%       & CmBERT-ptrn &  93.6 &  94.0 &  94.4 &  94.6 &  94.6 &  94.8 &  95.0 &  94.9 \\
%       & SpaCy NER &  88.6 &  90.4 &  90.7 &  91.7 &  91.9 &  92.8 &  93.3 &  93.4 \\
%\end{tabular}
%\end{table}

\begin{figure}[ht!]
	   \center{\includegraphics[width=\textwidth]
	       {figs/eval-ner-exp1.pdf}}
	  \caption{\label{fig:f1-vs-trainsize} Metrics measured on the fine-tuned models CmBERT, CmBERT+ptrn and SpaCy NER in experiment 1 for 8 training sets of increasing sizes. As SpaCy NER is outperformed by both BERT models from the start, CmBERT slightly surpasses the pre-trained version when the number of training examples becomes large. Values for precision and recall are available in the appendix.\bertrand{TODO!}}
\end{figure}
